---
title: "JCaguioa_Livesession3assignment"
author: "Joseph Caguioa"
date: "11/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1: GitHub Cloning
#### Use Git to clone https://github.com/caesar0301/awesome-public-datasets. Show code, commented out, for creating a new directory, navigating to the appropriate directory, and cloning the repository.

```{r GitHubClone}
# mkdir desktop/SMUDS/MSDS6306_DDS/Local_Repos
# cd desktop/SMUDS/MSDS6306_DDS/Local_Repos
# git clone https://github.com/awesomedata/awesome-public-datasets.git
```

------

# Question 2: Data Summary
#### *Part a*: In R, use read.table() or read.csv() to assign titanic.csv to df.

```{r TitanicRead}
# titanic.csv.zip extracted to same folder in unshown step.
setwd("/Users/josephcaguioa/Desktop/SMUDS/MSDS6306_DDS/Local_Repos/awesome-public-datasets/Datasets")
df <- read.csv("titanic.csv", header = TRUE, sep = ",")
str(df)
```

#### *Part b*: Output the gender counts. Plot their frequencies.

```{r TitanicGender}
# Output counts of females and males.
Gender <- df$Sex
FemaleCount <- sum(Gender == "female")
MaleCount <- sum(Gender == "male")
print(paste("There were", FemaleCount, "females and", MaleCount, "males aboard the Titanic.", sep = " "))

# Plot female and male frequencies.
Genders <- plot(Gender,
           main = "People Aboard the Titanic, by Gender",
           xlab = "Genders", ylab = "Number of People",
           ylim = c(0,600),
           col=c("lightcoral","lightblue")
           )
```

#### *Part c*: Use an *apply* function to obtain the means of Age, Fare, and Survival.

```{r TitanicApply, warning=FALSE}
# Use an *apply function to obtain means.
MeansList <- sapply(df, mean, na.rm=TRUE)

# Obtain indices on relevant columns for output.
AFSIndices <- vector("integer", 3)
AFSIndices[1] <- which(colnames(df) == "Survived")
AFSIndices[2] <- which(colnames(df) == "Age")
AFSIndices[3] <- which(colnames(df) == "Fare")
print(round(MeansList[AFSIndices], 2))
```

------

# Question 3: Function Building
#### *Parts a-d*: Create helper function(x){}. 

```{r HelpFunction}
SleepDataFunc <- function(x) {

  # Part a: Create objects for median Age, minimum and maximum Duration of sleep, and mean and standard deviation of RSES.
  MedianAge <- median(x$Age, na.rm = TRUE)
  MinSleep <- min(x$Duration, na.rm = TRUE)
  MaxSleep <- max(x$Duration, na.rm = TRUE)
  MeanRses <- mean(x$RSES, na.rm = TRUE)
  SdRes <- sd(x$RSES, na.rm = TRUE)
  
  # Part b: Create report data frame.
  report <- data.frame(MedianAge,
                       MeanRses / 5,
                       SdRes / 5,
                       MaxSleep - MinSleep)
  
  # Part c: Change column names.
  names(report) <- c("MedianAge",
                     "SelfEsteem",
                     "SE_SD",
                     "DurationRange")
  
  # Part d: Round report.
  round(report, 2)
}
```

#### *Part e*: Run function on sleep_data_01.csv.

```{r SleepRun}
SleepUrl <- "http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv"
sleep_df <- read.csv(SleepUrl, header = TRUE, sep = ",")
SleepDataFunc(sleep_df)
```

------

# Question 4: FiveThirtyEight Data
#### *Part a*: Install fivethirtyeight.

```{r FiveThirtyEightInstall, warning=FALSE}
install.packages("fivethirtyeight", repos = "https://github.com/rudeboybert/fivethirtyeight/")
library("fivethirtyeight")
```

#### *Part b*: Get 22nd data set.

```{r TheTwentySecond}
data(package = "fivethirtyeight")
df <- college_recent_grads
```

#### *Part c*: Write URL.

```{r TheTwentySecondUrl}
# http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```

#### *Part d*: Give dimensions and column names.

```{r TwentySecondStructure}
dim(df)
names(df)
```

------

# Question 5: Data Summary
#### *Part a*: Get column names and number of columns.

```{r ColumnInformation}
colnames(df)
ncol(df)
```

#### *Part b*: Count unique major_category, assign to major_count.

```{r MajorCategories, message = FALSE}
library(dplyr)
major_count <- df %>% group_by(major_category) %>%
  summarise(frequency = length(major_category))
print(major_count)
```

#### *Part c*: Make barplot of major_count.

```{r MajorCountPlot, fig.align = "center"}
barplot(major_count$frequency,
        names.arg = major_count$major_category,
        horiz = TRUE, axes = TRUE, las=2,
        xlim = c(0,30), xlab = "Frequency", ylab = "Categories",
        main = "Number of Majors by Category",
        col = "steelblue")
```

#### *Part d*: Write data to csv without row labels.

```{r FiveThirtyEightCsv}
write.csv(df, file = "fivethirtyeight_recentgrads.csv", row.names = FALSE)
```

------

# Question 6: Codebook
#### *Parts a-d*: Create GitHub MSDS homework with README.md and post 'fivethirtyeight' associated documents.

```{r Codebook}

# GitHub repo found here: https://github.com/J-Caguioa/DSHomework.git

```
