---
title: "JCaguioa_LiveSession5Assignment"
author: "Joseph Caguioa"
date: "12/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1: Data Munging

```{r munge2016, message=FALSE}

library(tidyr)
library(dplyr)

# Part a: Import file, rename columns.
df <- read.csv2("./yob2016.txt", header = FALSE, check.names = FALSE)
colnames(df) <- c("Name", "Gender", "Births_2016")

# Part b: Display summary and structure.
summary(df)
str(df)

# Part c: Find erroneous name.
aWeirdName <- grep("y{3}", df$Name, value = TRUE)
print(aWeirdName)

# Part d: Begone, Fionayyy!
y2016 <- data.frame(df[df$Name != 'Fionayyy', ])

```

***

## Question 2: Data Merging

```{r merge}

# Part a: Import file, rename columns.
y2015 <- read.csv("./yob2015.txt", header = FALSE, check.names = FALSE)
colnames(y2015) <- c("Name", "Gender", "Births_2015")

# Part b: Display last 10 rows.
tail(y2015, 10)

```

Rows 33054-33063, the last ten of the "y2015" dataframe, all describe _"popular"_ boy names beginning with the letter "Z" that were each given to five newborns in 2015. I have never heard of any of these, and will be very shocked if I meet one in my lifetime. ~~It seems these unique-sounding names are not quite so unique, seeing as each one was assigned to five babies in 2015 alone.~~

```{r mergeContd2}

# Demonstrate non-unique problem with Name column.
str(y2015)
str(y2016)
y2016[y2016$Name == "Joseph", ]

```

Attempting to merge as is gives errors because merging requires columns of unique values. When looking at str(y2015) and str(y2016), the number of levels in First_Name is less than the number of observations, indicating that some names are repeated (given as a unisex name). To work around this, Name and Gender are momentarily united to create a column of unique name-gender observations, and then separated again.

```{r mergeContd3}

# Part c: Merge dataframes, only keeping names with data for both years.
# Temporarily unite Name and Gender for merging purposes.
y2015_2 <- unite(y2015, NameGender, c(Name, Gender), remove = TRUE)
y2016_2 <- unite(y2016, NameGender, c(Name, Gender), remove = TRUE)
both_years <- merge(y2015_2, y2016_2, by = "NameGender", all = FALSE)

# Separate Name and Gender columns again.
final <- separate(both_years, col = "NameGender", into = c("Name", "Gender"), sep = "_", remove = TRUE)

# Demonstrate that no columns contain NA values after merge.
colSums(is.na(final))

```

***

## Question 3: Data Summary

```{r summary}

# Part a: Create Total column.
final$Total <- final$Births_2015 + final$Births_2016
popular_named_kids <- sum(final$Total)
cat("In 2015 and 2016 combined, ", popular_named_kids, " babies received popular names!", sep = "")

# Part b: Top ten most popular names.
sorted_final <- final %>% arrange(-Total)
head(sorted_final, 10)

# Part c: Top ten most popular girl names.
sorted_girls <- sorted_final[sorted_final$Gender %in% "F", ]
top_ten_girls <- sorted_girls[c(1:10), c(1,5)]
top_ten_girls

# Part d: Output top ten most popular girl names. Placed into same Unit_5_Homework folder.
write.csv(top_ten_girls, "./top_girl_names.csv", row.names = FALSE)

```

***

## Question 4: GitHub

Please see the following repository on my GitHub: https://github.com/J-Caguioa/DSHomework. The Unit_5_Homework folder should contain this RMarkdown file, the knitted HTML, the referenced yob2015.txt and yob2016.txt raw documents, the top_girl_names.csv output, a codebook, and a custom css file aptly named custom_html.css.
