---
title: "Live Session 12 Assignment: Time Series"
author: "Joseph Caguioa"
date: "2/4/2019"
output:
  html_document:
    theme: default
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
---

```{r setup, include=T, message=F, warning=F, results="hide"}
knitr::opts_chunk$set(echo = TRUE)

# Load packages.
package_list <- c("tidyverse", "fpp2", "dygraphs", "xts", "MuMIn", "kableExtra")
lapply(package_list, library, character.only=T)

```

Note that no data is normalized in this document.

# Q1: Brief Financial Data

```{r q1, fig.align='center', message=F, warning=F}

# 1a: Pull DAX index from EuStockMarkets.
help(EuStockMarkets)
dax <- EuStockMarkets[,"DAX"]

# 1b: Plot DAX.
plot(dax, main="Daily closing price of DAX, 1991-1998",
     xlab="Year", ylab="Daily Closing Price",
     col="blue")
abline(v=1997, col="red")

# 1c: Decompose DAX into multiplicative model.
plot(decompose(dax, type="multiplicative"), 
     cex.lab=0.75, cex.axis=0.75, col="blue")
abline(v=1997, col="red")

```

# Q2: Temperature Data

```{r q2, fig.align='center', message=F, warning=F}

# 2a: Load maxtemp dataset.
help(maxtemp)
autoplot(maxtemp)

# 2b: Subset to data after 1990.
maxtemp_1990 <- window(maxtemp, start=1990)

# 2c: Use SES to predict next 5 years.
maxtemp_1990_ses <- ses(maxtemp_1990, h=5)
plot(maxtemp_1990_ses, sub="Melbourne's Maximum Temperatures: Forecast with SES",
     xlab="Year", ylab="Maximum Temperature (°C)")
lines(fitted(maxtemp_1990_ses), col="blue")
legend("topleft", legend=c("Historical Data", "Predicted Values"), col=c("black", "blue"),
       box.lty=0, lty=1)

# 2d: Use damped Holt's linear trend to predict next 5 years.
maxtemp_1990_holt <- holt(maxtemp_1990, initial="optimal", damped=T, h=5)
plot(maxtemp_1990_holt, sub="Melbourne's Maximum Temperatures: Forecast with Holt",
     xlab="Year", ylab="Maximum Temperature (°C)")
lines(fitted(maxtemp_1990_holt), col="blue")
legend("topleft", legend=c("Historical Data", "Predicted Values"), col=c("black", "blue"),
       box.lty=0, lty=1)

# 2e: Compare AICc of ses() and holt().
maxtemp_aic <- rbind(c("AIC", AIC(maxtemp_1990_ses$model), AIC(maxtemp_1990_holt$model)),
                     c("AICc", AICc(maxtemp_1990_ses$model), AICc(maxtemp_1990_holt$model)))
kable(maxtemp_aic, col.names=c("Criterion", "SES", "Holt"), row.names=F,
      caption="Selection Criteria", table.attr="style='width:30%;'") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

When considering the Akaike Information Criterion (AIC) or its alternative, the second-order Akaike Information Criterion (AICc), the selection method involves minimizing the selection criterion. In this case, the AIC/AICc of simple exponential smoothing (SES; 140.48/141.53) is lower than that of the damped Holt model (141.38/145.58). The ses() model is better.

# Q3: The Wands Choose the Wizard

```{r q3, fig.align='center', message=F, warning=F}

# 3a: Load data frames.
gregorovitch <- read_csv("Unit11TimeSeries_Gregorovitch.csv", col_names=c("Date", "Wands"))
ollivander <- read_csv("Unit11TimeSeries_Ollivander.csv", col_names=c("Date", "Wands"))

# 3b: Convert to Date class.
gregorovitch$Date <- as.Date(gregorovitch$Date, format="%m/%d/%Y")
ollivander$Date <- as.Date(ollivander$Date, format="%m/%d/%Y")

# 3c: Convert data frames to xts, order by Date.
gregorovitch_xts <- xts(gregorovitch$Wands, order.by=gregorovitch$Date)
ollivander_xts <- xts(ollivander$Wands, order.by=ollivander$Date)

# 3d: Bind xts objects and create dygraph.
wands_xts <- merge(gregorovitch_xts, ollivander_xts)
dygraph(wands_xts, main="Wands Sold from 1970-2016", xlab="Year", ylab="Wands Sold") %>%
  dySeries(label=c("Gregorovitch")) %>% dySeries(label=c("Ollivander")) %>%
  dyOptions(stackedGraph=T, colors=RColorBrewer::brewer.pal(2, "Set2")) %>%
  dyRangeSelector() %>%
  dyShading(from="1995-01-01", to="1999-01-01", color="#FF1C1C") %>%
  dyHighlight() %>%
  dyAxis("y", valueRange=c(0,4000)) %>%
  dyLegend(width=400)

```